---
title: "Lab5"
author: "Frank"
date: "2023-03-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Setting the working directory to the folder where this file is.
```{r}
# use package "here"
library(tidyverse)
```

Loading the data files.
```{r}
provenOilReserveWEurope <- read.csv("provenOilReserveWEurope.csv", header=T, stringsAsFactors=T)
channels <- read.csv("channels.csv", header=T, stringsAsFactors=T)
consumptionTransposed <- read.csv("consumptionTransposed.csv", header=T, stringsAsFactors=T)
consumptionBySource <- read.csv("consumptionBySource.csv", header=T, stringsAsFactors=T)
```

## Plotting channels
```{r}
p <- ggplot(channels, aes(n))
p <- p+geom_line(aes(y = electric, colour= "electric"))
p <- p+geom_line(aes(y = length, colour ="length"))
p <- p+geom_line(aes(y = area, colour= "area"))
p<- p+scale_colour_hue("channels")
p
```
### Adding better labels
```{r}
p <- ggplot(channels, aes(n))
p <- p+geom_line(aes(y = electric, colour= "electric"))
p <- p+geom_line(aes(y = length, colour ="length"))
p <- p+geom_line(aes(y = area, colour= "area"))
p <- p+scale_colour_hue("channels")
p <-p+labs(x= "Size of stimuli", y= "Amount of sensation")
p
```
### Adding better legends
```{r}
p <- ggplot(channels, aes(n))
p <- p+geom_line(aes(y = electric, colour = "electric current"))
p <- p+geom_line(aes(y = length, colour = "object’s length"))
p <- p+geom_line(aes(y = area, colour = "object’s area"))
p <- p+scale_colour_hue("channels")
p<-p+labs(x= "size of stimuli", y= "size of sensation")
p
```
### Changing y axis limits  to see detail for bottom part of plot. However;
A warning will appear as not all data for electric is displayed (some of it falls outside the y limits).
```{r}
p <- ggplot(channels, aes(n))
p <- p+geom_line(aes(y = electric, colour = "electric current"))
p <- p+geom_line(aes(y = length, colour = "object’s length"))
p <- p+geom_line(aes(y = area, colour = "object’s area"))
p <- p+scale_colour_hue("channels")
p<-p+labs(x= "size of stimuli", y= "size of sensation")
p <- p+ylim(0,50) 
p
```
### Removing electric
```{r}
# make a copy
channels2 <- channels
channels2$electric <-NULL 
p <- ggplot(channels2, aes(n))
p <- p+geom_line(aes(y = length, colour = "object’s length"))
p <- p+geom_line(aes(y = area, colour = "object’s area"))
p <- p+scale_colour_hue("channels")
p<-p+labs(x= "size of stimuli", y= "size of sensation")
p
```
### Removing all rows where the value for electric is over 50.
```{r}
channels2 <- channels[channels$electric < 50,]
p <- ggplot(channels2, aes(n))
p <- p+geom_line(aes(y = electric, colour = "electric current"))
p <- p+geom_line(aes(y = length, colour = "object’s length"))
p <- p+geom_line(aes(y = area, colour = "object’s area"))
p <- p+scale_colour_hue("channels")
p<-p+labs(x= "size of stimuli", y= "size of sensation")
p
```
## Facets
### Side by side plots
```{r}
p<- ggplot(provenOilReserveWEurope, aes(Year))
p <- p+geom_line(aes(y = MT.Barrels))
p<- p + facet_grid(. ~ Country)
p
```
```{r}
p<- ggplot(provenOilReserveWEurope, aes(Year))
p <- p+geom_line(aes(y = MT.Barrels, colour = Country))
p<- p + facet_grid(. ~ Country)
p
```
### One plot under the other
```{r}
p<- ggplot(provenOilReserveWEurope, aes(Year))
p <- p+geom_line(aes(y = MT.Barrels))
p<- p + facet_grid(Country ~ .)
p
```
#### Adding colour
```{r}
P <- ggplot(provenOilReserveWEurope, aes(Year))
p <- p+geom_line(aes(y = MT.Barrels, colour=Country))
p <- p + facet_grid(Country ~ .)
p
```
#### Free y axis (different y scale for different plots)
```{r}
P <- ggplot(provenOilReserveWEurope, aes(Year))
p <- p+geom_line(aes(y = MT.Barrels, colour=Country))
p <- p + facet_grid(Country ~ ., scales="free_y")
p
```
#### Free axis (both x and y axis)
```{r}
p<- ggplot(provenOilReserveWEurope, aes(Year)) 
p <- p+geom_line(aes(y = MT.Barrels, colour=Country,linetype=Country))
p<- p + facet_grid(Country ~ ., scales = "free")
p
```
#### Using a point geom
```{r}
p <- ggplot(provenOilReserveWEurope, aes(Year))
p <- p+geom_point(aes(y = MT.Barrels, colour=Country))
p <- p + facet_grid(Country ~ ., scales="free_y")
p
```
#### Using both point and line geoms
```{r}
p <- ggplot(provenOilReserveWEurope, aes(Year))
p <- p+geom_point(aes(y = MT.Barrels, colour=Country))
p <- p+geom_line(aes(y = MT.Barrels, colour=Country))
p<- p + facet_grid(Country ~ ., scales="free_y")
p
```
#### Using free space - space used for plot depends on data shown.
```{r}
p <- ggplot(provenOilReserveWEurope, aes(Year))
p <- p+geom_point(aes(y = MT.Barrels, colour=Country))
p <- p+geom_line(aes(y = MT.Barrels, colour=Country))
p <- p + facet_grid(Country ~ ., scales = "free", space="free_y")
p
```
First a solution with colours for the various energy sources chosen by default.
```{r}
p <- ggplot(consumptionTransposed, aes(Year))
p <- p + geom_point(aes(y=Oil, colour="Oil") )
p <- p + geom_point(aes(y=Gas, colour="Gas") )
 p <- p + geom_point(aes(y=Coal, colour="Coal") )
p <- p + geom_point(aes(y=Renewables, colour="Renewables (all kinds)"))
p <- p + labs(x= "Year", y= "Millions of tons of oil equivalent", title="Energy consumed")
p <- p+scale_colour_hue("Energy source")
p
```
```{r}
# Assigning the colours I want to use to the labels in the legand area
plotColours <- c("Oil" = "brown", "Gas"="yellow", "Coal"= "black", "Renewables (all kinds)"="green")

p <- ggplot(consumptionTransposed, aes(Year))
p <- p + geom_point(aes(y=Oil, colour="Oil") )
p <- p + geom_point(aes(y=Gas, colour="Gas") )
 p <- p + geom_point(aes(y=Coal, colour="Coal") )
p <- p + geom_point(aes(y=Renewables, colour="Renewables (all kinds)"))
p <- p + labs(x= "Year", y= "Millions of tons of oil equivalent", title="Energy consumed",colour="Energy source")
p <- p+ scale_colour_manual(values=plotColours)
p
```
```{r}
p <- ggplot(consumptionBySource, aes(Year))
p <- p+geom_line(aes(y = MTOE, colour=Source))
p <- p + facet_grid(Source ~ ., scales="free_y")
p <- p + labs(x= "Year", y= "Millions of tons of oil equivalent", title="Energy consumed", colour="Energy source")
p
```
```{r}
p <- ggplot(consumptionBySource, aes(Year))
p <- p+geom_point(aes(y = MTOE, colour=Source), shape= 24)
p <- p + facet_grid(Source ~ ., scales="free_y")
p <- p + labs(x= "Year", y= "Millions of tons of oil equivalent", title="Energy consumed", colour="Energy source")
p
```
### Exercise 3 - alternative if we want the triangles to be filled
```{r}
p <- ggplot(consumptionBySource, aes(Year))
p <- p+geom_point(aes(y = MTOE, colour=Source, fill=Source), shape= 24)
p <- p + facet_grid(Source ~ ., scales="free_y")
p <- p + labs(x= "Year", y= "Millions of tons of oil equivalent", title="Energy consumed")
p
```

