---
title: "Week4-Lab4"
author: "Frank"
date: "2023-03-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Setting the working directory to the folder where this file is.
```{r Warning = F}
rm(list=ls())
```
Loading  library (package)  tidyverse
```{r Warning = F}
library(tidyverse)
```
Loading all the data.
```{r}
consumptionNoOil <- read.csv("consumptionNoOil.csv", header=T, stringsAsFactors=T)
ProvenOilReservesZerosT <- read.csv("ProvenOilReservesZerosT.csv", header=T, stringsAsFactors=T)
ProvenOilReserveWEurope <- read.csv("ProvenOilReserveWEurope.csv", header=T, stringsAsFactors=T)
channels <- read.csv("channels.csv", header=T, stringsAsFactors=T)
universities <- read.csv("universities.csv", header=T, stringsAsFactors=T)
importsExportsAmerica<- read.csv("importsExportsAmerica.csv", header=T, stringsAsFactors=T)
projectTeam <- read.csv("projectTeam.csv", header=T, stringsAsFactors=T)
```
#Using a text geom
Instead of points  text labels are used in this plot.
```{r}
p <- ggplot(importsExportsAmerica, aes(crude.imports,crude.exports, label=country))
p <- p+geom_text()
p
```
## Bar plots
### Stacking
```{r}
p <- ggplot(projectTeam, aes(site, fill = experience))
p <- p + geom_bar(position = "stack")
p
```
### Dodging
```{r}
p <- ggplot(projectTeam, aes(site, fill = experience))
p <- p + geom_bar(position = "dodge")
p
```
### Filling
```{r}
p <- ggplot(projectTeam, aes(site, fill=experience))
p <- p + geom_bar(position = "fill")
p
```
## Identity
Note that in this case some bars are hidden behind other bars so this is not very useful.
```{r}
p <- ggplot(projectTeam, aes(site, fill = experience))
p <- p + geom_bar(position = "identity")
p
```
## Adjusting axis
The following plot has no axis adjustments.
```{r}
p <- ggplot(consumptionNoOil, aes(Year,MTOE, colour=Source))
p <- p+geom_point()
p
```
### Adjusting the x axis
```{r}
p <- ggplot(consumptionNoOil, aes(Year,MTOE, colour=Source)) 
p <- p + xlim(2010,2035)
p <- p+geom_point()
p
```
## Adding titles, subtitles,  x and y axis labels and levend label
```{r}
p <- ggplot(consumptionNoOil, aes(Year,MTOE, colour=Source))
p <- p+geom_point()
p<- p+ labs(x="Year",
            y= "Millions of tons of oil equivalent", 
            title ="Energy consumption by source", 
            subtitle = "Oil is excluded",
            colour = "Energy source")

p
```
## Changing colours
```{r}
colours <-c(coal="black", gas="yellow2", renewables="green")
p <- ggplot(consumptionNoOil, aes(Year, MTOE, colour=Source)) 
p <- p +geom_point()
p<-p+labs(x= "Year", 
          y= "Millions of tons of oil equivalent", 
          title = "Energy consumed", 
          colour="Energy source")
p<- p + scale_colour_manual(values=colours)
p
```
## Adding lines
```{r}
p <- ggplot(consumptionNoOil, aes(Year,MTOE, colour=Source)) 
p <- p+geom_point()
p <- p + geom_vline(aes(xintercept = Year ), linetype = "longdash", colour="green")
p
```
```{r}
p <- ggplot(consumptionNoOil, aes(Year,MTOE, colour=Source)) 
p <- p+geom_point()
p <- p + geom_vline(aes(xintercept = 2015),  linetype = "longdash", colour="green")
p
```
```{r}
p <- ggplot(consumptionNoOil, aes(Year,MTOE)) 
p <- p+geom_point()
p <- p + geom_hline(aes(yintercept = 100), linetype = "dotdash", colour="green")
p
```
```{r}
p <- ggplot(consumptionNoOil, aes(Year,MTOE, colour=Source)) 
p <- p+geom_point()
p <- p + geom_hline(aes(yintercept = 100), linetype = "dotted", colour="darkgreen")
p <- p + geom_vline(aes(xintercept = 2015),  linetype = "twodash", colour="blue")
p
```
```{r}
p <- ggplot(consumptionNoOil, aes(Year,MTOE, colour=Source)) 
p <- p+geom_point()
p <- p + geom_hline(aes(yintercept = 100), linetype = "dashed", colour="green")
p <- p + geom_vline(aes(xintercept = 2015), linetype = "dashed", colour="blue")
p
```
## Pie chart
```{r}
pieData <- projectTeam |> count(site, name="Number")

names(pieData) <- c("site","numberStaff")
## order data according to the site (important for placing labels later on)

pieData <- arrange(pieData,desc(site))
# create new column with position for label
pieData <- pieData |>  mutate(positionLabel = cumsum(numberStaff) - 0.5* numberStaff)
# creating plot
p <- ggplot(pieData,  aes(x="", y= numberStaff, fill = site))
p <- p + geom_bar(width = 1, stat="identity")
p <- p + coord_polar("y", start=0)

# Changing the colour scheme
p <- p +   scale_fill_manual(values = c("red","blue", "lightblue")) 

# removing outer "ring"
p <- p + theme_void()
p
```
```{r}
p <- ggplot(ProvenOilReservesZerosT, aes(MTBarrels, Denmark))
p <- p + geom_point()
p <- p + labs(x="Year", y="Million tons oil equivalent" , title="Proven Oil Reserves in Denmark") 
p
```
This produces far too many vertical lines!
```{r}
p <- ggplot(ProvenOilReservesZerosT, aes(MTBarrels, Denmark))
p <- p + geom_point()
p <- p + labs(x="Year", y="Million tons oil equivalent" , title="Proven Oil Reserves in Denmark") 
p <- p + geom_vline(aes(xintercept = MTBarrels),linetype = "longdash")
p
```
###solution - with only a couple of  vertical lines!
```{r}
p <- ggplot(ProvenOilReservesZerosT, aes(MTBarrels, Denmark))
p <- p + geom_point()
p <- p + labs(x="Year", y="Million tons oil equivalent" , title="Proven Oil Reserves in Denmark") 

# lines put at 1990 and 2005. These are just examples. 
p <- p + geom_vline(aes(xintercept = 1990))
p <- p + geom_vline(aes(xintercept = 2005))
p
```
```{r}
p <- ggplot(ProvenOilReservesZerosT, aes(MTBarrels, Italy))
p <- p + geom_line(colour="orange", size=2)
p <- p + labs(x="Year", y="Million tons oil equivalent" , title="Proven Oil Reserves in Italy") 
p <- p + geom_hline(aes(yintercept = 0.9))
p
```
In the solution below note that using colour to emphasize size may not be the best option.
```{r}
p <- ggplot(ProvenOilReservesZerosT, aes(MTBarrels, United.Kingdom, colour=United.Kingdom))
p <- p + geom_point()
p <- p + labs(x="Year", y="Million tons oil equivalent" , title="Proven Oil Reserves in the UK", colour="Size of Reserves") 
### horizontal blue line added at 5.
p <- p + geom_hline(aes(yintercept = 5), colour="blue")
p
```
```{r}
p <- ggplot(ProvenOilReserveWEurope, aes(Year, MT.Barrels, group=Country))
p <- p + geom_line(aes(colour=factor(Country)))
p <- p + labs(x="Year", y="Million tons barrels" , title="Proven Oil Reserves in Western Europe", colour= "Country") 
p
```
```{r}
p <- ggplot(channels, aes(length, area, size=n))
p <- p + geom_point() 
p <- p + labs(x="length", y="area" , title="Perception - length vs. area", size= "Size of stimuli") 
p
```
```{r}
p <- ggplot(channels, aes(length, area, size=n))
p <- p + geom_point()
p <- p + labs(x="length", y="area" , title="Perception - length vs. area", size= "Size of stimuli") 
p <- p + ylim(0,5)
p
```
```{r}
p <- ggplot(channels, aes(length, area, size=n,))
p <- p + geom_point(colour = "limegreen")
p <- p + labs(x="length", y="area" , title="Perception - length vs. area", size= "Size of stimuli") 
p <- p + ylim(0,5)
p
```
### stacking
```{r}
p <- ggplot(universities, aes(status, fill = school))
p <- p + geom_bar(position = "stack")
p
```
```{r}
p <- ggplot(universities, aes(status, fill = school))
p <- p + geom_bar(position = "dodge")
p
```
#### Filling
```{r}
p <- ggplot(universities, aes(status, fill = school))
p <- p + geom_bar(position = "fill")
p
```
#### Identity 
This does not make any sense as some data is "hidden" behind other data.
```{r}
p <- ggplot(universities, aes(status, fill = school))
p <- p + geom_bar(position = "identity")
p
```
```{r}
pieData <- projectTeam |> count(site, name="Number")
universities2 <- universities |> filter(status %in% c("student", "both"))

# Counting the number of rows per school.
#pieData <- count(universities2$school)
names(pieData) <- c("school","numberStudents")

#Oordering data according to the school name (important for placing labels later on)
pieData <- arrange(pieData,desc(school))

# Creating a new column called positionLabel with position for label
pieData <-  mutate(pieData, positionLabel = cumsum(pieData$numberStudents) - 0.5*pieData$numberStudents)

# creating pie chart
p <- ggplot(pieData,  aes(x="", y= numberStudents, fill = school))
p <- p + geom_bar(width = 1, stat="identity")
p <- p + coord_polar("y", start=0)

# Adding text labels
p <- p + geom_text(aes(y = positionLabel, label = numberStudents))

# Changing the colour scheme
p <- p +   scale_fill_manual(values = c("wheat", "lightblue", "cornflowerblue", "blue1")) 
# removing outer "ring"
p <- p + theme_void()

p
```

